---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document: 
    keep_md: true
---


## Loading and preprocessing the data :sheep:
Load the data and add a new variable called datetime with the full date and time of the interval.
```{r cache=TRUE, results="hide"}
activity<-read.csv("activity.csv")

require(stringr)
require(plyr)
activity$interval<-str_sub(paste("000",activity$interval,sep=""),-4,-1)
activity<-mutate(activity,datetime=as.POSIXct(paste(activity$date,activity$interval),format=("%Y-%m-%d %H%M")))

```

## What is mean total number of steps taken per day?
```{r total_steps}
total_per_day<-tapply(activity$steps,activity$date,sum)
mean_total<-round(mean(total_per_day,na.rm=TRUE),digits=2)
median_total<-median(total_per_day,na.rm=TRUE)
hist(total_per_day,breaks="FD")
```
  
The mean steps per day is `r round(mean_total)` steps and the median of total steps per day is `r median_total` steps.

## What is the average daily activity pattern?
```{r mean per interval}
mean_per_interval<-tapply(activity$steps,activity$interval,mean,na.rm=TRUE)
times<-unique(strftime(activity$datetime,format="%H:%M:%S"))
max_time<-times[mean_per_interval==max(mean_per_interval)]
plot(as.data.frame(times),mean_per_interval,xlab="time of day",ylab="mean steps per 5 minutes",type="l")
```

The interval with the highest mean of steps is at `r max_time`.

## Imputing missing values
```{r}
missing_values<-is.na(activity$steps)
steps_filled<-activity$steps
long_mean_per_interval<-rep(mean_per_interval,length.out=length(steps_filled))
fill_data<-function(x) steps_filled[x]<-round(long_mean_per_interval[x])
replaced<-sapply(which(missing_values),fill_data)
activity_filled<-cbind(steps_filled,activity[2:4])

total_per_day_filled<-tapply(activity_filled$steps_filled,activity_filled$date,sum)
mean_total_filled<-mean(total_per_day_filled,na.rm=TRUE)
median_total_filled<-median(total_per_day_filled,na.rm=TRUE)
hist(total_per_day_filled,breaks="FD")
```

The total number of missing values is `r sum(missing_values)`. Missing values were replaced by the rounded mean of the corresponding interval. The mean and median of the new data set are then `r mean_total_filled` and `r median_total_filled`, identical to the original data set.

## Are there differences in activity patterns between weekdays and weekends?
```{r day_plot, results="hide"}
require(plyr)
require(reshape2)
activity<-mutate(activity,weekday=weekdays(datetime))
days<-cbind(c(rep("weekday",5),rep("weekend",2)),unique(activity$weekday))
activity<-mutate(activity,day=as.factor(mapvalues(weekday,days[,2],days[,1])))


mean_per_interval_per_day<-tapply(activity$steps,list(activity$interval,activity$day),mean,na.rm=TRUE)

mean_per_interval_per_day<-melt(mean_per_interval_per_day)
names(mean_per_interval_per_day)<-c("interval","day","mean_steps")

require(lattice)

xyplot(mean_steps ~ interval | day , data = mean_per_interval_per_day,layout=c(1,2),type="l")

```